\section{Working with remote machines}

\subsection{SSH}
\begin{itemize}
\item \textbf{Usage}\\
\texttt{\$ ssh host}\\
\texttt{\$ password:}\\

\item \textbf{Examples of host}\\
\texttt{192.162.82.120}\\
\texttt{bioclust.myuniversity.edu}\\
\texttt{darwin@192.162.82.120}\\
\texttt{darwin@bio.univ.edu}\\

\item \textbf{Options} \\

\texttt{-v} verbose. Verbosity can be increased by: \texttt{-vv} or \texttt{-vvv}. \\
\texttt{-p} port. Eg: \texttt{ssh -p 5043 cdarwin@bio.univ.edu}\\
\qquad Default port is \texttt{22}

 
\item \textbf{Using alias:} To use alias create the file \texttt{\~{}/.ssh/config} and store server as info as below. 
\texttt{Host bio\_serv}\\
\qquad \texttt{HostName 190.512.171.29}\\
\qquad \texttt{User cdarwin}\\
\qquad \texttt{Port 50434}\\
Also applies for \texttt{Rsync and scp}\\

\item \textbf{SSH keys}: SSH key to connect without password. Eg:\\
\texttt{\$ ssh-keygen -b 2048}\\
This command request the following:\\
\begin{itemize}
\item File to save the key. By default this is:\\
\texttt{/Users/username/.ssh/id\_rsa}
NOTE: This file is the private key.\\
\item Passphrase:\\
Not necessary but good to use.\\
\end{itemize}
Private key: \texttt{~/.ssh/id\_rsa}\\
Public key: \texttt{~/.ssh/id\_rsa.pub}\\

\texttt{\$ chmod 400 id\_rsa \# restrict access to private key}\\
\texttt{\$ ssh-add} \\

\end{itemize}


\subsection{Establishing a server}
Use "Open SSH": https://help.ubuntu.com/lts/serverguide/openssh-server.html\\

My IP address: \texttt{hostname -I}\\
List of logins to the server: \texttt{sudo less /var/log/auth.log}\\

\subsection{nohup}
\texttt{nohup} runs program un-interupted. \\
The execution continues even when the terminal is closed or connection to remote machine is lost.\\
\begin{itemize}
\item Just add \texttt{nohup} just before the command. \\
Eg: \texttt{nohup prog1}
\item Usually nohup is in the background.\\
Eg: \texttt{nohup prog1 \&}
\end{itemize}

\textcolor{red}{Unfinished: Tmux.}

\vfill \null
\columnbreak
